Improving \textsl{Gaia} parallax precision \\ with a data-driven model of stars

### 2017-06-14 paper/paper.tex
figuresize
apsis

The \gaia\ Mission (\citealt{gaia16}) will soon deliver more than a billion
stellar parallax measurements.
Only a small fraction (but large number) of these
measurements will be both precise and purely astrometric.
\gaia\ will use astrometric parallaxes to determine the distances of the more
precise stars, and then calibrate spectrophotometric models.
These spectrophotometric models, along with \gaia's on-board
low-resolution $BP/RP$ spectrophotometry, are used to provide
parallax estimates to less precise stars.
The full stack required for these parallax inferences is complex.
It involves modeling the stars, as well as the dust, in the Milky Way,
and the response of the telescope itself.
It might be possible to eliminate some of this complexity using data driven models which don't rely on physical models.

@@

\begin{figure}
\centering
\includegraphics[width=0.75\textwidth]{deltaCDF.pdf}
\caption{Fractional Change in Variance: The cumulative distribution function of the natural log of the fractional variance change when comparing our posterior to the \gaia\ catalogue. Negative numbers represent stars which have a posterior that is narrower than the \gaia\ catalogue, which is true for a large fraction of the stars. We've drawn two dashed lines. One represents the median star $(y=0.5)$ which has a fractional variation of $x=-0.3=ln(1/1.2^2)$, corresponding to an increase in precision by a factor of 1.2. Another represents the fraction of stars $(y=0.14)$ which have a fractional variance greater than $x=-1.4=ln(1/2^2)$, corresponding to an increase in precision by a factor of 2.}
\label{fig:deltaCDF}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{deltaParallax.pdf}
\caption{The difference between the expectation value of our parallax posteriors and the \gaia\ parallax measurement as a function of the \gaia\ uncertainty. The points are colored by the fractional change in the variance of the posterior PDF compared with the \gaia\ likelihood, the same as Figure \ref{fig:delta}. The points are evenly distributed around $y=0$ showing that there is little bias introduced by inferred parallax PDFs. The distribution gets wider with larger \gaia\ uncertainties due to those lower signal to noise measurements being more influenced by the prior. They also tend to have the largest fractional change in their variance (shown as darker points)}
\label{fig:deltaParallax}
\end{figure}

>>>>

The \gaia\ Mission (\citealt{gaia16}) will soon deliver more than a billion
stellar parallax measurements.
Only a small fraction (but large number) of these
measurements will be precise and purely astrometric.
\gaia\ will use astrometric parallaxes to determine the distances of the more
precise stars, and then calibrate spectrophotometric models (\citealt{apsis}).
These spectrophotometric models, along with \gaia's on-board
low-resolution $BP/RP$ spectrophotometry, can be used to provide
more precise parallax estimates for stars with poor astrometric parallax data.
The full stack required for these parallax inferences is complex.
It involves modeling the stars, as well as the dust, in the Milky Way,
and the response of the telescope itself.
It might be possible to eliminate some of this complexity using data driven models which don't rely on physical models.

@@

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{deltaCDF.pdf}
\caption{Fractional Change in Variance: The cumulative distribution function of the natural log of the fractional variance change when comparing our posterior to the \gaia\ catalogue. Negative numbers represent stars which have a posterior that is narrower than the \gaia\ catalogue, which is true for a large fraction of the stars. We've drawn two dashed lines. One represents the median star $(y=0.5)$ which has a fractional variation of $x=-0.3=ln(1/1.2^2)$, corresponding to an increase in precision by a factor of 1.2. Another represents the fraction of stars $(y=0.14)$ which have a fractional variance greater than $x=-1.4=ln(1/2^2)$, corresponding to an increase in precision by a factor of 2.}
\label{fig:deltaCDF}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{deltaParallax.pdf}
\caption{The difference between the expectation value of our parallax posteriors and the \gaia\ parallax measurement as a function of the \gaia\ uncertainty. The points are colored by the fractional change in the variance of the posterior PDF compared with the \gaia\ likelihood, the same as Figure \ref{fig:delta}. The points are evenly distributed around $y=0$ showing that there is little bias introduced by inferred parallax PDFs. The distribution gets wider with larger \gaia\ uncertainties due to those lower signal to noise measurements being more influenced by the prior. They also tend to have the largest fractional change in their variance (shown as darker points)}
\label{fig:deltaParallax}
\end{figure}
